Approximation as Differentiation

### 2025-06-09 main/approx-as-tangents.tex
Some blurb on CMon-enrichment.

\section{Approximations as Tangents}

We motivate our approach to \GPS by showing how to combine ideas from differential geometry and stable domain theory to reconstruct \GPS in a dentational setting. From this basis, we apply the CHAD framework of V{\'a}k{\'a}r \etal to obtain a denotational model of automatic differentiation that gives us an effective way to interpret programs as functions along with their forward and backwards approximation maps\bob{and we did it in Agda, so we can run it}.

>>>>

\section{Approximations as Tangents}
\label{sec:approx-as-tangents}

We motivate our approach to \GPS by showing how to combine ideas from differential geometry and stable domain theory to reconstruct \GPS in a denotational setting. From this basis, we apply the CHAD framework of V{\'a}k{\'a}r \etal to obtain a denotational model of automatic differentiation that gives us an effective way to interpret programs as functions along with their forward and backwards approximation maps\bob{and we did it in Agda, so we can run it}.

### 2025-06-09 main/first-order.tex
Some blurb on CMon-enrichment.

\subsection{Biproducts and Semi-Additive Categories}
\label{sec:biproducts}

@@

One might also like to allow the programmer to embed additional ``approximation points'' into composite
structures, by providing the language with a \emph{lifting} monad $\syn{L}$. Rather than the usual definition
of $\syn{list}\;\syn{nat}$ as $\mu\alpha.1 + \syn{nat} \times \alpha$, one could then define within the
language a type of ``approximation lists'' $\mu\alpha.\syn{L}\;(1 + \syn{L}\;\syn{nat} \times \alpha)$, along
with an approximating version of multiplication $*_{\text{appr}}: \syn{L}\;\syn{nat} \times \syn{L}\;\syn{nat}
\to \syn{L}\;\syn{nat}$ expressible as
\[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' =
0\;\syn{then}\;\tmReturn{0}\;\syn{else}\;\tmBind{y}{\tmFun{y'}{\tmReturn{x' * y'}}}}}\].

>>>>

\subsection{$\CMon$-Categories and Biproducts}
\label{sec:biproducts}

For differentiable functions, we can combine linear approximations by adding tangent vectors. For stable
functions, we can combine approximants by taking their meet. The additive structure in both cases form a
commutative monoid with an addition + and zero element 0. If this additive structure is also present in the
hom-sets $\cat{C}(X,Y)$ of a category, and composition preserves it, $\cat{C}$ is enriched in $\CMon$, the
category of commutative monoids.

@@

One might also like to allow the programmer to embed additional ``approximation points'' into composite
structures, by providing the language with a \emph{lifting} monad $\syn{L}$. Rather than the usual definition
of $\syn{list}\;\syn{nat}$ as $\mu\alpha.1 + \syn{nat} \times \alpha$, one could then define (within the
language) a type of ``approximation lists'' $\mu\alpha.\syn{L}\;(1 + \syn{L}\;\syn{nat} \times \alpha)$, along
with an approximating version of multiplication $*_{\text{appr}}: \syn{L}\;\syn{nat} \times \syn{L}\;\syn{nat}
\to \syn{L}\;\syn{nat}$ expressible as
\[x *_{\text{appr}} y = \tmBind{x}{\tmFun{x'}{\syn{if}\;x' =
0\;\syn{then}\;\tmReturn{0}\;\syn{else}\;\tmBind{y}{\tmFun{y'}{\tmReturn{x' * y'}}}}}\].
